\HeaderA{CairoFonts}{Set the fonts used for all Cairo graphics devices.}{CairoFonts}
\keyword{device}{CairoFonts}
\begin{Description}\relax
\code{CairoFonts} initializes the fonts used for Cairo graphics devices.
\end{Description}
\begin{Usage}
\begin{verbatim}
CairoFonts(
        regular="Helvetica:style=Regular",
        bold="Helvetica:style=Bold",
        italic="Helvetica:style=Italic",
        bolditalic="Helvetica:style=Bold Italic,BoldItalic",
        symbol="Symbol"
)
\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{regular}] character; fontconfig pattern for the 'plain text' font.
\item[\code{bold}] character; fontconfig pattern for the 'bold face' font.
\item[\code{italic}] character; fontconfig pattern for the 'italic' font.
\item[\code{bolditalic}] character; fontconfig pattern for the 'bold italic' font.
\item[\code{symbol}] character; fontconfig pattern for the 'symbol' font.
\end{ldescription}
\end{Arguments}
\begin{Details}\relax
This function sets the fonts for Cairo graphics devices globally; previously opened
Cairo graphics devices will also use these fonts. The argument names correspond to 
the five values of the graphical parameter 'font', i.e. regular is 1, bold is 2, italic 
is 3, etc.

For an explanation of fontconfig patterns, see \code{\LinkA{CairoFontMatch}{CairoFontMatch}}.
\end{Details}
\begin{Section}{Known issues}
\Itemize{
\item This function is only available when the cairo graphics library is configured
with FreeType and Fontcofig support.
\item R math symbols are supported,but require a
"Symbol" font with the Adobe symbol encoding. At the time of this release
the authors are unaware of a free version of this font, however they do exist on Windows
and Mac OS X.
}
\end{Section}
\begin{SeeAlso}\relax
\code{\LinkA{CairoFontMatch}{CairoFontMatch}},
\code{\LinkA{Cairo}{Cairo}},
\code{\LinkA{par}{par}},
\end{SeeAlso}
\begin{Examples}
\begin{ExampleCode}
## Not run: 
#
# The following fontconfig patterns define the free truetype fonts 
# available in the debian package 'ttf-freefont'.
# 
# Freesans is very similar to Helvetica
CairoFonts(
        regular="FreeSans:style=Medium",
        bold="FreeSans:style=Bold",
        italic="FreeSans:style=Oblique",
        bolditalic="FreeSans:style=BoldOblique"
)
## End(Not run)\end{ExampleCode}
\end{Examples}

