\HeaderA{prediction-class}{Class "prediction"}{prediction.Rdash.class}
\keyword{classes}{prediction-class}
\begin{Description}\relax
Object to encapsulate numerical predictions together with the
corresponding true class labels, optionally collecting predictions and
labels for several cross-validation or bootstrapping runs.
\end{Description}
\begin{Section}{Objects from the Class}
Objects can be created by using the
\code{prediction} function.
\end{Section}
\begin{Section}{Slots}
\describe{
\item[\code{predictions}:] A list, in which each element is a vector of
predictions (the list has length \textgreater{} 1 for x-validation data.)
\item[\code{labels}:] Analogously, a list in which each element is a
vector of true class labels.
\item[\code{cutoffs}:] A list in which each element is a vector of
all necessary cutoffs. Each cutoff vector consists of the
predicted scores (duplicates removed), in descending order.
\item[\code{fp}:] A list in which each element is a vector of the number (not
the rate!) of false positives induced by the cutoffs given in the
corresponding 'cutoffs' list entry.
\item[\code{tp}:] As fp, but for true positives.
\item[\code{tn}:] As fp, but for true negatives.
\item[\code{fn}:] As fp, but for false negatives.
\item[\code{n.pos}:] A list in which each element contains the number of
positive samples in the given x-validation run.
\item[\code{n.neg}:] As n.pos, but for negative samples.
\item[\code{n.pos.pred}:] A list in which each element is a vector
of the number of samples predicted as positive at the cutoffs
given in the corresponding 'cutoffs' entry.
\item[\code{n.neg.pred}:] As n.pos.pred, but for negatively
predicted samples.
}
\end{Section}
\begin{Note}\relax
Every \code{prediction} object contains information about the 2x2
contingency table consisting of tp,tn,fp, and fn, along with the
marginal sums n.pos,n.neg,n.pos.pred,n.neg.pred, because these form
the basis for many derived performance measures.
\end{Note}
\begin{Author}\relax
Tobias Sing \email{tobias.sing@mpi-sb.mpg.de},
Oliver Sander \email{osander@mpi-sb.mpg.de}
\end{Author}
\begin{References}\relax
A detailed list of references can be found on the ROCR
homepage at \url{http://rocr.bioinf.mpi-sb.mpg.de}.
\end{References}
\begin{SeeAlso}\relax
\code{\LinkA{prediction}{prediction}}, \code{\LinkA{performance}{performance}},
\code{\LinkA{performance-class}{performance.Rdash.class}}, \code{\LinkA{plot.performance}{plot.performance}}
\end{SeeAlso}

